Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-02 ms.tex
analytics

@@

Continuity of the entropy derivative at the stagnation radius where $v = 0$ requires that the temperature at this location be given by (eq.~[\ref{eq:first_law}])
\begin{align}
T|_{r_{s}} \simeq \frac{\gamma-1}{\gamma}\frac{\mu m_p \tilde{v}_{w}^{2}}{2k} \approx 3.7
\times 10^6\ \, v_{500}^2 \,\,{\rm K},
\label{eq:Tanalytic}
\end{align}
{\bf AG missing correction here for black hole velocity dispersion}
where $v_{500} \equiv \tilde{v}_{w}/(500$ km s$^{-1}$).  Hydrostatic equilibrium furthermore determines the value of the stagnation radius (Appendix \ref{app:rs}, eq.~\ref{eq:rs2main})
\begin{align}
\rs=\frac{G \Mbh}{n \vw^{2}|_{\rs}}\left(\frac{9-\Gamma}{2} \frac{M_{\star}|_{\rs}}{\Mbh} +\frac{7}{2}\right),
\label{eq:stag_analytic}
\end{align}
where $n \equiv -d{\rm ln}\rho/d{\rm ln\,r}|_{r_{\rm s}}$ is the density power-law slope at $r = r_{\rm s}$.

If the SMBH dominates the gravitational potential at the stagnation radius, then $v_{w} \gtrsim \sigma$ and hence the stagnation radius resides well inside the sphere of influence ($r_{\rm s} < r_{\rm soi}$; eq.~[\ref{eq:rsoi}]).  In this limit, equation (\ref{eq:stag_analytic}) simplifies to
\begin{eqnarray}
  &\rs \underset{v_w \gtrsim \sigma}\approx \frac{7}{2}\frac{G \Mbh}{n \tilde{v}_{w}^{2}} \approx \begin{cases}
    12
 \, \pc \,\, \Mbheight v_{500}^{-2}\, \pyear& \text{core} \\
    6
 \, \pc \,\, \Mbheight v_{500}^{-2} \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:stag_simple}
\end{eqnarray}
{\bf AG: missing a minus n here...}
where for the numerical estimate we adopt values of $n \approx 1$ and $n \approx 0.5$ for core ($\Gamma = 0.1$) and cusp ($\Gamma = 0.8$) galaxies, respectively, so chosen to match the results of our numerical calculations (Fig.~\ref{fig:stag}).  This expression is similar to that obtained by \citet{Volonteri+11} on more heuristic grounds (their eq.~6).  {\bf BDM: all analytic expressions have to be updated in core case to reflect this larger rs}

@@

The accretion rate onto the SMBH is given by the integrated mass loss rate interior to the stagnation radius (eq.~[\ref{eq:q}]), 
\begin{eqnarray}
  \dot{M} &=& 4\pi \int_{0}^{r_{s}}q r^{2}dr = \frac{\eta \Menc|_{r_{\rm s}}}{\th} \nonumber \\
&\approx&
  \begin{cases}
    2.9 \times 10^{-5} M_{\bullet,8}^{1.76}
    v_{500}^{-3.8}  \eta_{0.02} \Msun \, \pyear& \text{core} \\
    5.3 \times 10^{-5} M_{\bullet,8}^{1.48} 
    v_{500}^{-2.4}  \eta_{0.02} \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:mdot_analytic}
\end{eqnarray}
where we have assumed $v_{w} \gg \sigma$ by adopting equation (\ref{eq:stag_simple}) for $r_s$.  The resulting Eddington ratio is given by 
\begin{eqnarray}
\frac{\dot{M}}{\dot{M}_{\rm edd}} &\approx&
  \begin{cases}
    1.3 \times 10^{-5} M_{\bullet,8}^{0.76}
    v_{500}^{-3.8}  \eta_{0.02}   & \text{core}, \\
    2.4 \times 10^{-5} \Mbheight^{0.48} 
    v_{500}^{-2.4}  \eta_{0.02}   & \text{cusp}, 
  \end{cases}
  \label{eq:eddr_analytic}
\end{eqnarray}
where $\dot{M}_{\rm edd} = 2.6M_{\bullet,8}M_{\odot}$ yr$^{-1}$  is the Eddington accretion rate, assuming a radiative efficiency of ten per cent.  Note the sensitive inverse dependence of the accretion rate on the wind heating rate.  
%To reproduce the estimated values $\dot{M}/\dot{M}_{\rm edd} \lesssim 10^{-3}$ in low-luminosity AGN or quiescent galaxies, we thus require $v_{w} \gtrsim 150M_{\bullet,8}^{0.5}$

The gas density at the stagnation radius $\rho|_{r_{\rm s}}$ is more challenging to estimate accurately.  Using an alternative estimate of $\dot{M}$ as the gaseous mass within the stagnation radius divided by the free-fall time $t_{\rm ff}|_{r_{\rm s}} = (r_{\rm s}^{3}/GM_{\bullet})^{1/2}$,
\begin{align}
  &\dot{M}\sim\frac{(4 \pi/3) \rs^3 \rho|_{r_{\rm s}}}{\tff|_{r_{\rm s}}},
  \label{eq:mdot_gas}
\end{align}
 in conjuction with equations (\ref{eq:mdot_analytic}) and (\ref{eq:rs_simple}), one derives
\begin{align}
  \rho|_{r_{\rm s}}\approx
  \begin{cases}
    5 \times 10^{-26} \Mbheight^{-0.2} v_{500}^{-0.8}  \eta_{0.02} \,
    \, {\rm g \, cm^{-3}}& \text{core},\\
    9 \times 10^{-26}  \Mbheight^{-0.5} v_{500}^{0.6}  \eta_{0.02} \,\, {\rm g \,cm^{-3}} & \text{cusp}.
  \end{cases}
  \label{eq:rhors}
\end{align}

@@

One observes a similarity between equation (\ref{eq:mdot_gas}) and the Bondi formula (eq.~[\ref{eq:bondi}]) if $r_{\rm B}$ is replaced by $\rs$.  Indeed, when the stagnation radius resides interior to the sphere of influence we find that (eq.~[\ref{eq:stag_simple}])
\begin{align}
  \rs\approx\frac{7}{2}\frac{G \Mbh}{n \vwO^2} \approx \frac{7}{6n}r_{\rm B},
  \label{eq:rs_simple}
\end{align}
where the second equality makes use of equation (\ref{eq:Tanalytic}).

@@

Our analytic derviations neglect the effects of heat conduction, an assumption we now check.  The ratio of the magnitude of the conductive heating rate to the external heating rate at the stagnation radius is given by
\begin{align}
  \left.\frac{\nabla\cdot(\kappa \nabla T)}{q v_{\rm
w}^{2}/2}\right|_{r_{\rm s}} &\sim \frac{2t_{\rm h}\kappa_0
T|_{r_{\rm s}}^{7/2}}{r_{\rm s}^{2}\eta \rho_{\star}|_{r_{\rm s}} \tilde{v}_{w}^{2}(1+\phi)}
\nonumber \\ &\sim {\rm min}
  \begin{cases}
  10 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.8} v_{500}^{6.8} &  \text{unsaturated (core)}\\
 10 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.5} v_{500}^{5.4} &  \text{unsaturated (cusp)}\\
  1.0(\phi/0.1)/(2-\Gamma) & \text{saturated},
  \end{cases}
 \label{eq:conduction}
\end{align}
where the second equality makes use of equations (\ref{eq:Tanalytic}), (\ref{eq:stag_simple}), and we have approximated $\nabla^{2} \sim 1/r_{\rm s}^{2}$, $\nabla \sim 1/r_{\rm s}$.  We have approximated the stellar density profile as (eq.~[\ref{eq:rhostar}])
\begin{eqnarray}
  \rho_{\star}|_{r_{\rm s}} &\simeq& \frac{M_{\bullet}(2-\Gamma)}{4\pi r_{\rm soi}^{3}}\left(\frac{r_{\rm s}}{r_{\rm soi}}\right)^{-1-\Gamma} \nonumber \\
 &\approx& \begin{cases}
    9.4 \times 10^{-19}M_{\bullet,8}^{-1.2}v_{500}^{2.2}\,{\rm g\,cm^{-3}}
    & \text{core} \\
    1.1 \times 10^{-18} M_{\bullet,8}^{-1.5}v_{500}^{3.6}
    \,{\rm g\,cm^{-3}}  & \text{cusp}, 
  \end{cases}
  \label{eq:rhostarrs}
\end{eqnarray}
where the stagnation radius is assume to reside well inside the Nuker break radius.  Equation (\ref{eq:conduction}) shows that even in the case of saturated conduction, the neglect of heat conduction results in at most an order unity correction for expected values of the saturation parameter $\phi < 0.1$, justifying its neglect in our analysis.  Numerical experiments including conductivity confirm this ($\S\ref{sec:numerical}$).

@@

Figure~\ref{fig:profiles} shows profiles of the density $\rho(r)$,
temperature $T(r)$, and radial velocity $|v(r)|$ (expressed in units of the sound speed), for the cusp
($\Gamma=0.8$) solutions within our grid.  As expected, the gas
density increases towards the SMBH $\rho\propto r^{-n}$ with
$n\simeq1$, i.e. shallower than the $-3/2$ power law for
Bondi accretion. This power law behavior does not extend
through all radii, however, as the gas density profile has a break coincident
with the location of the break in the stellar light profile ($\rb=100$
pc). The temperature profile is relatively flat at large radii, but
increases as $\propto 1/r^{k}$ interior to the sphere of influence,
where $k\lesssim 1$, somewhat shallower than expected for virialized gas within the black hole sphere of influence.
The inwardly directed velocity increases towards the hole with a profile that is somewhat steeper than the local free-fall velocity $v \propto v_{\rm ff}\propto r^{-1/2}$.  The flow near the stagnation radius is subsonic, but becomes supersonic at two critical points, one located at small radii, $r \lesssim 0.1r_{\rm s}$, and the other located at large radii, $r \gtrsim 10^{3}r_{\rm s}$.  In many cases the supersonic transition is instigated by the transition to a steeper stellar density prorfile exterior to the outer Nuker break radius, making our solutions near $r_{\rm s}$ also potentially sensitive to the break radius value.
% {\bf BDM: I made this up - in general we need
%   more explanation of what the plots are actually showing, with some
%   physical explanation.}
% AG -- things are not quite approaching Bondi on the inner grid. I
% don't have a good physical explanation yet...

Figure~\ref{fig:stag} shows our results for the stagnation radius $r_{\rm s}$, expressed in ratio to sphere of influence radius $r_{\rm soi}$, with different colors showing different values of $v_{w}$.  Cusp and core galaxies are marked with square and triangles, respectively.  Shown for comparison are our analytic results (eq.~[\ref{eq:rs2main}]) with solid and dashed lines for cusp and core galaxies, respectively, calculated assuming $n = 1$ and $n= 0.5$, respectively.  Our analytic estimates are found to provide accurate fits to the numerical results for $v_{\rm w} \gtrsim \sigma$, but the assumptions leading to these estimates become invalid for low $v_{\rm w}$ as the stagnation radius diverges to large radii.  For $v_{w} \gg \sigma$ the ratio $\rs/\rsoi$ does not depend sensitively on break radius $r_{b}$, but the latter has a greater influence when $v_{w} \lesssim \sigma$.  The three green squares which are vertically aligned at the same value of $v_{w}/\sigma$ (close to where the analytic estimate diverges) correspond to models with the same black hole mass ($\Mbh = 10^7 \Msun$) and heating rate ($v_w = 300$ km s$^{-1}$) but different values of the break radius: $r_{b} = $ 50, 100, and 400 pc (from top to bottom).  As $r_{\rm b} \rightarrow r_{s}$, the effects of the break radius is to the push the stagnation radius inwards, decreasing the SMBH accretion rate.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{rs.eps}
  \caption{\label{fig:stag} \emph{Top panel:} Stagnation radius
    $r_{s}$ in units of the sphere of influence radius $r_{\rm
      soi}$ (eq.~[\ref{eq:rsoi}]) for galaxies in our sample as a
    function of the ratio of the effective wind heating rate $v_{w}$
    to the stellar velocity dispersion $\sigma$.  Green, orange, and
    blue symbols correspond to different values of $v_{w} =$ 300, 600, and 1200 km s$^{-1}$,
    respectively.  Squares correspond to cusp galaxies ($\Gamma = 0.8$), while
    triangles correspond to cores ($\Gamma = 0.1$). The black curves correspond to the
    analytic prediction from equation (\ref{eq:rs2main}), with solid and dashed curves calculated for parameters $\{\Gamma=0.8, n=1\}$ and $\{\Gamma=0.1,n=0.5\}$, respectively. }
\end{figure}

@@

The importance of Compton heating can be estimated by assuming the SMBH
radiates with a luminosity $L = \epsilon_{\rm rad} \dot{M}c^{2}$, where
$\epsilon_{\rm rad}$ is the radiative efficiency and where $\dot{M}$ is
estimated from equation (\ref{eq:mdot_analytic}).  Then using
equations (\ref{eq:rs_simple}), (\ref{eq:mdot_analytic}),
(\ref{eq:rhors}), (\ref{eq:rhostarrs}) we calculate that from equation
(\ref{eq:vw_eff}) that the effective heating rate at the stagnation radius is given by
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 29 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.38}v_{500}^{-1.4}\,{\rm
km\,s^{-1}} &, \text{core}\\ 50 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.24}v_{500}^{-0.7}\,{\rm
km\,s^{-1}} &, \text{cusp},
  \end{cases}
  \label{eq:vC}
\end{align} where $T_{C,9} = T_{C}/10^{9}$ K and $\epsilon_{-2} =
\epsilon_{\rm rad}/0.01 \sim 1$ is the maximum radiative efficiency
for an advective accretion flow expected to produce high $T_{\rm C}$
emission (e.g.~\citealt{Narayan&Yi95}).  We caveat that, unlike stellar wind heating, Compton heating depends on radius, scaling as $v_{w}^{\rm C}(r) \propto (n^{2}\xi/\rho_{\star})^{1/2} \propto r^{(\Gamma-n-1)/2}$, i.e. $\propto r^{-0.6}$ and $\propto r^{-0.7}$ for core ($\Gamma = 0.8$; n = 1) and cusp ($\Gamma = 0.1$; n = 0.5) galaxies, respectively.  The assumption of our model that the heating parameter be constant with radius is not satisfied, but this radial variation is sufficiently weak that it should not significantly alter our conclusions.

>>>>

\newcommand{\densSlope}{\nu}

@@

Continuity of the entropy derivative at the stagnation radius where $v = 0$ requires that the temperature at this location be given by (eq.~[\ref{eq:first_law}])
\begin{align}
T|_{r_{s}} \simeq \frac{\gamma-1}{\gamma}\frac{\mu m_p \tilde{v}_{w}^{2}}{2k} \approx 3.7
\times 10^6\ \, v_{500}^2 \,\,{\rm K},
\label{eq:Tanalytic}
\end{align}
{\bf AG missing correction here for black hole velocity dispersion}
where $v_{500} \equiv \tilde{v}_{w}/(500$ km s$^{-1}$).  Hydrostatic equilibrium furthermore determines the value of the stagnation radius (Appendix \ref{app:rs}, eq.~\ref{eq:rs2main})
\begin{align}
  \rs=\frac{G \Mbh}{\densSlope
    v_w^{2}|_{\rs}}\left[\left(\frac{9-\Gamma}{2} -\densSlope\right)
    \frac{\Mstar|_{\rs}}{\Mbh} +\frac{7}{2}-\densSlope\right].
\label{eq:stag_analytic}
\end{align}
where $\densSlope \equiv -d{\rm ln}\rho/d{\rm ln\,r}|_{r_{\rm s}}$ is the
density power-law slope at $r = r_{\rm s}$. We find that

\begin{align}
\densSlope \simeq \frac{1}{6} \left(4 \Gamma+3\right).
\label{eq:densSlope}
\end{align}

If the SMBH dominates the gravitational potential at the stagnation
radius, then $v_{w} \gtrsim \sigma$ and hence the stagnation radius
resides well inside the sphere of influence ($r_{\rm s} < r_{\rm
  soi}$; eq.~[\ref{eq:rsoi}]).  In this limit, equation
(\ref{eq:stag_analytic}) simplifies to
\begin{eqnarray}
  &\rs \underset{v_w \gtrsim \sigma}\approx \left(\frac{7}{2}-\densSlope\right)\frac{G \Mbh}{\densSlope v_w^2} \approx \begin{cases}
    9
 \, \pc \,\, \Mbheight v_{500}^{-2}\, \pyear& \text{core} \\
    4
 \, \pc \,\, \Mbheight v_{500}^{-2} \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:stag_simple}
\end{eqnarray}

where we have used equation~\eqref{eq:densSlope} to estimate
$\densSlope$: $\densSlope\approx 1$ and $\densSlope \approx 0.6$ for
core ($\Gamma = 0.1$) and cusp ($\Gamma = 0.8$) galaxies.  This
expression is similar to that obtained by \citet{Volonteri+11} on more
heuristic grounds (their eq.~6).  {\bf BDM: all analytic expressions
  have to be updated in core case to reflect this larger rs}

@@

The accretion rate onto the SMBH is given by the integrated mass loss rate interior to the stagnation radius (eq.~[\ref{eq:q}]), 
\begin{eqnarray}
  \dot{M} &=& 4\pi \int_{0}^{r_{s}}q r^{2}dr = \frac{\eta \Menc|_{r_{\rm s}}}{\th} \nonumber \\
&\approx&
  \begin{cases}
    6.1 \times 10^{-5} M_{\bullet,8}^{1.76}
    v_{500}^{-3.8}  \eta_{0.02} \Msun \, \pyear& \text{core} \\
    3.3 \times 10^{-5} M_{\bullet,8}^{1.48} 
    v_{500}^{-2.4}  \eta_{0.02} \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:mdot_analytic}
\end{eqnarray}
where we have assumed $v_{w} \gg \sigma$ by adopting equation (\ref{eq:stag_simple}) for $r_s$.  The resulting Eddington ratio is given by 
\begin{eqnarray}
\frac{\dot{M}}{\dot{M}_{\rm edd}} &\approx&
  \begin{cases}
    2.8 \times 10^{-5} M_{\bullet,8}^{0.76}
    v_{500}^{-3.8}  \eta_{0.02}   & \text{core}, \\
    1.5 \times 10^{-5} \Mbheight^{0.48} 
    v_{500}^{-2.4}  \eta_{0.02}   & \text{cusp}, 
  \end{cases}
  \label{eq:eddr_analytic}
\end{eqnarray}
where $\dot{M}_{\rm edd} = 2.6M_{\bullet,8}M_{\odot}$ yr$^{-1}$  is the Eddington accretion rate, assuming a radiative efficiency of ten per cent.  Note the sensitive inverse dependence of the accretion rate on the wind heating rate.  
%To reproduce the estimated values $\dot{M}/\dot{M}_{\rm edd} \lesssim 10^{-3}$ in low-luminosity AGN or quiescent galaxies, we thus require $v_{w} \gtrsim 150M_{\bullet,8}^{0.5}$

The gas density at the stagnation radius $\rho|_{r_{\rm s}}$ is more challenging to estimate accurately.  Using an alternative estimate of $\dot{M}$ as the gaseous mass within the stagnation radius divided by the free-fall time $t_{\rm ff}|_{r_{\rm s}} = (r_{\rm s}^{3}/GM_{\bullet})^{1/2}$,
\begin{align}
  &\dot{M}\sim\frac{(4 \pi/3) \rs^3 \rho|_{r_{\rm s}}}{\tff|_{r_{\rm s}}},
  \label{eq:mdot_gas}
\end{align}
 in conjuction with equations (\ref{eq:mdot_analytic}) and (\ref{eq:rs_simple}), one derives
\begin{align}
  \rho|_{r_{\rm s}}\approx
  \begin{cases}
    6 \times 10^{-26} \Mbheight^{-0.2} v_{500}^{-0.8}  \eta_{0.02} \,
    \, {\rm g \, cm^{-3}}& \text{core},\\
    10 \times 10^{-26}  \Mbheight^{-0.5} v_{500}^{0.6}  \eta_{0.02} \,\, {\rm g \,cm^{-3}} & \text{cusp}.
  \end{cases}
  \label{eq:rhors}
\end{align}

@@

One observes a similarity between equation (\ref{eq:mdot_gas}) and the Bondi formula (eq.~[\ref{eq:bondi}]) if $r_{\rm B}$ is replaced by $\rs$.  Indeed, when the stagnation radius resides interior to the sphere of influence we find that (eq.~[\ref{eq:stag_simple}])
\begin{align}
  \rs\approx\frac{7}{2}\frac{G \Mbh}{\densSlope \vw^2} \approx
  \frac{7}{6 \densSlope}r_{\rm B},
  \label{eq:rs_simple}
\end{align}
where the second equality makes use of equation (\ref{eq:Tanalytic}).

@@

Our analytic derviations neglect the effects of heat conduction, an assumption we now check.  The ratio of the magnitude of the conductive heating rate to the external heating rate at the stagnation radius is given by
\begin{align}
  \left.\frac{\nabla\cdot(\kappa \nabla T)}{q v_{\rm
w}^{2}/2}\right|_{r_{\rm s}} &\sim \frac{2t_{\rm h}\kappa_0
T|_{r_{\rm s}}^{7/2}}{r_{\rm s}^{2}\eta \rho_{\star}|_{r_{\rm s}} \tilde{v}_{w}^{2}(1+\phi)}
\nonumber \\ &\sim {\rm min}
  \begin{cases}
  10 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.8} v_{500}^{6.8} &  \text{unsaturated (core)}\\
 10 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.5} v_{500}^{5.4} &  \text{unsaturated (cusp)}\\
  1.0(\phi/0.1)/(2-\Gamma) & \text{saturated},
  \end{cases}
 \label{eq:conduction}
\end{align}
where the second equality makes use of equations (\ref{eq:Tanalytic}), (\ref{eq:stag_simple}), and we have approximated $\nabla^{2} \sim 1/r_{\rm s}^{2}$, $\nabla \sim 1/r_{\rm s}$.  We have approximated the stellar density profile as (eq.~[\ref{eq:rhostar}])
\begin{eqnarray}
  \rho_{\star}|_{r_{\rm s}} &\simeq& \frac{M_{\bullet}(2-\Gamma)}{4\pi r_{\rm soi}^{3}}\left(\frac{r_{\rm s}}{r_{\rm soi}}\right)^{-1-\Gamma} \nonumber \\
 &\approx& \begin{cases}
    6.1 \times 10^{-19}M_{\bullet,8}^{-1.2}v_{500}^{2.2}\,{\rm g\,cm^{-3}}
    & \text{core} \\
    2.1 \times 10^{-18} M_{\bullet,8}^{-1.5}v_{500}^{3.6}
    \,{\rm g\,cm^{-3}}  & \text{cusp}, 
  \end{cases}
  \label{eq:rhostarrs}
\end{eqnarray}
where the stagnation radius is assume to reside well inside the Nuker break radius.  Equation (\ref{eq:conduction}) shows that even in the case of saturated conduction, the neglect of heat conduction results in at most an order unity correction for expected values of the saturation parameter $\phi < 0.1$, justifying its neglect in our analysis.  Numerical experiments including conductivity confirm this ($\S\ref{sec:numerical}$).

@@

Figure~\ref{fig:profiles} shows profiles of the density $\rho(r)$,
temperature $T(r)$, and radial velocity $|v(r)|$ (expressed in units of the sound speed), for the cusp
($\Gamma=0.8$) solutions within our grid.  As expected, the gas
density increases towards the SMBH $\rho\propto r^{-\densSlope}$ with
$\densSlope\simeq1$, i.e. shallower than the $-3/2$ power law for
Bondi accretion. This power law behavior does not extend
through all radii, however, as the gas density profile has a break coincident
with the location of the break in the stellar light profile ($\rb=100$
pc). The temperature profile is relatively flat at large radii, but
increases as $\propto 1/r^{k}$ interior to the sphere of influence,
where $k\lesssim 1$, somewhat shallower than expected for virialized gas within the black hole sphere of influence.
The inwardly directed velocity increases towards the hole with a profile that is somewhat steeper than the local free-fall velocity $v \propto v_{\rm ff}\propto r^{-1/2}$.  The flow near the stagnation radius is subsonic, but becomes supersonic at two critical points, one located at small radii, $r \lesssim 0.1r_{\rm s}$, and the other located at large radii, $r \gtrsim 10^{3}r_{\rm s}$.  In many cases the supersonic transition is instigated by the transition to a steeper stellar density prorfile exterior to the outer Nuker break radius, making our solutions near $r_{\rm s}$ also potentially sensitive to the break radius value.
% {\bf BDM: I made this up - in general we need
%   more explanation of what the plots are actually showing, with some
%   physical explanation.}
% AG -- things are not quite approaching Bondi on the inner grid. I
% don't have a good physical explanation yet...

Figure~\ref{fig:stag} shows our results for the stagnation radius
$r_{\rm s}$, expressed in ratio to sphere of influence radius $r_{\rm
  soi}$, with different colors showing different values of $v_{w}$.
Cusp and core galaxies are marked with square and triangles,
respectively.  Shown for comparison are our analytic results
(eq.~[\ref{eq:rs2main}]) with solid and dashed lines for cusp and core
galaxies, respectively, calculated assuming $\densSlope = 1$ and
$\densSlope= 0.5$, respectively.  Our analytic estimates are found to
provide accurate fits to the numerical results for $v_{\rm w} \gtrsim
\sigma$, but the assumptions leading to these estimates become invalid
for low $v_{\rm w}$ as the stagnation radius diverges to large radii.
For $v_{w} \gg \sigma$ the ratio $\rs/\rsoi$ does not depend
sensitively on break radius $r_{b}$, but the latter has a greater
influence when $v_{w} \lesssim \sigma$.  The three green squares which
are vertically aligned at the same value of $v_{w}/\sigma$ (close to
where the analytic estimate diverges) correspond to models with the
same black hole mass ($\Mbh = 10^7 \Msun$) and heating rate ($v_w =
300$ km s$^{-1}$) but different values of the break radius: $r_{b} = $
50, 100, and 400 pc (from top to bottom).  As $r_{\rm b} \rightarrow
r_{s}$, the effects of the break radius is to the push the stagnation
radius inwards, decreasing the SMBH accretion rate.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{rs.eps}
  \caption{\label{fig:stag} \emph{Top panel:} Stagnation radius
    $r_{s}$ in units of the sphere of influence radius $r_{\rm soi}$
    (eq.~[\ref{eq:rsoi}]) for galaxies in our sample as a function of
    the ratio of the effective wind heating rate $v_{w}$ to the
    stellar velocity dispersion $\sigma$.  Green, orange, and blue
    symbols correspond to different values of $v_{w} =$ 300, 600, and
    1200 km s$^{-1}$, respectively.  Squares correspond to cusp
    galaxies ($\Gamma = 0.8$), while triangles correspond to cores
    ($\Gamma = 0.1$). The black curves correspond to the analytic
    prediction from equation (\ref{eq:rs2main}), with solid and dashed
    curves calculated for parameters $\{\Gamma=0.8, \densSlope=1\}$ and
    $\{\Gamma=0.1,\densSlope=0.5\}$, respectively. }
\end{figure}

@@

The importance of Compton heating can be estimated by assuming the SMBH
radiates with a luminosity $L = \epsilon_{\rm rad} \dot{M}c^{2}$, where
$\epsilon_{\rm rad}$ is the radiative efficiency and where $\dot{M}$ is
estimated from equation (\ref{eq:mdot_analytic}).  Then using
equations (\ref{eq:rs_simple}), (\ref{eq:mdot_analytic}),
(\ref{eq:rhors}), (\ref{eq:rhostarrs}) we calculate that from equation
(\ref{eq:vw_eff}) that the effective heating rate at the stagnation radius is given by
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 29 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.38}v_{500}^{-1.4}\,{\rm
km\,s^{-1}} &, \text{core}\\ 50 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.24}v_{500}^{-0.7}\,{\rm
km\,s^{-1}} &, \text{cusp},
  \end{cases}
  \label{eq:vC}
\end{align} where $T_{C,9} = T_{C}/10^{9}$ K and $\epsilon_{-2} =
\epsilon_{\rm rad}/0.01 \sim 1$ is the maximum radiative efficiency
for an advective accretion flow expected to produce high $T_{\rm C}$
emission (e.g.~\citealt{Narayan&Yi95}).  We caveat that, unlike
stellar wind heating, Compton heating depends on radius, scaling as
$v_{w}^{\rm C}(r) \propto (n^{2}\xi/\rho_{\star})^{1/2} \propto
r^{(\Gamma-\densSlope-1)/2}$, i.e. $\propto r^{-0.6}$ and $\propto r^{-0.7}$
for core ($\Gamma = 0.8$; $\densSlope = 1$) and cusp ($\Gamma = 0.1$;
$\densSlope = 0.5$) galaxies, respectively.  The assumption of our model that the heating
parameter be constant with radius is not satisfied, but this radial
variation is sufficiently weak that it should not significantly alter
our conclusions.
